---
title: "Image Analysis"
output: html_notebook
---
```{r}
# Clear RStudio ---------------------------------------------------------
rm(list=ls(all=TRUE)) # clear workspace
graphics.off() # closes all graphics

```

```{r packages function}
packages<-function(x){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}
```


```{r get data}
data <- read.csv("fluorescence.csv",header=TRUE)
data
```



```{r}
# Preprocessing
# Visualize original data
packages(reshape)
packages(ggplot2)
multiplehist <- melt(data[,c(2:ncol(data))])
ggplot(multiplehist,aes(x = value)) + 
    facet_wrap(~variable,scales = "free") + 
    geom_histogram(col="black", 
                 fill="grey")
# Transformation
packages(caret)
packages(e1071)
trans = preProcess(data[,c(2:ncol(data))],c("YeoJohnson", "center", "scale"))
kos_trans = data.frame(trans = predict(trans, data))
kos_trans
# Visualize again
multiplehist <- melt(kos_trans)
ggplot(multiplehist,aes(x = value)) + 
    facet_wrap(~variable,scales = "free") + 
    geom_histogram(col="black", 
                 fill="grey")

```

```{r}
# Build model for original data
  dep.vars <- cbind(data[,c(2:ncol(data))])
  manova_fit <- manova(as.matrix(dep.vars) ~ data$Sample)
  summary(manova_fit, test="Wilks")
  cat("\n")
  summary(aov(as.matrix(dep.vars) ~ data$Sample))
  
  #LDA
  packages(MASS)
  lda <- lda(formula = Sample ~ ., 
           data = data
           )
  
  #Ordinal plot
  packages(devtools)
  packages(ggord)
  
  ord <- lda
  
  p <- ggord(ord, data$Sample)
  p #Note: needs to be cleaned up. Legend is messed up: shows two groupings.



```
```{r}
# Build model for transformed data
dep.vars <- cbind(kos_trans[,c(2:ncol(kos_trans))])
manova_fit <- manova(as.matrix(dep.vars) ~ kos_trans$trans.Sample)
summary(manova_fit, test="Wilks")
cat("\n")
summary(aov(as.matrix(dep.vars) ~ kos_trans$trans.Sample))
#LDA
  packages(MASS)
  lda <- lda(formula = trans.Sample ~ ., 
           data = kos_trans
           )
  
  #Ordinal plot
  packages(devtools)
  packages(ggord)
  
  ord <- lda
  
  p <- ggord(ord, kos_trans$trans.Sample)
  p #Note: needs to be cleaned up. Legend is messed up: shows two groupings.

```
This data looks decent. But how do we know we are getting representative samples of our biofilm? Let's check out EBImage (https://www.r-bloggers.com/r-image-analysis-using-ebimage/)
```{r EBImage}
# source("http://bioconductor.org/biocLite.R")
# biocLite()
# biocLite("EBImage")
require(EBImage)
```

```{r load image}
img <- readImage("Z-Stack GFP 15.tif")*2^16

display(normalize(img),method="raster",all=TRUE)


color_img <- rgbImage(green=img)
display(normalize(color_img),all=TRUE,metho="raster")
```
Great, so we can see our image. Can we measure integrated density like in ImageJ?
```{r img analysis}
img
dim(img)
hist(img)
range(img)

mean(img) # Mean pixel intensity
sd(img)
sum(img) # Integrated density
```
Here comes the real question. How many times do we have to sample a section of this biofilm to get an accurate picture of the mean/integrated density?
```{r}
#https://stackoverflow.com/questions/24299171/function-to-split-a-matrix-into-sub-matrices-in-r
matsplitter<-function(M, r, c) {
    rg <- (row(M)-1)%/%r+1
    cg <- (col(M)-1)%/%c+1
    rci <- (rg-1)*max(cg) + cg
    N <- prod(dim(M))/r/c
    cv <- unlist(lapply(1:N, function(x) M[rci==x]))
    dim(cv)<-c(r,c,N)
    cv
} 

split_img <- matsplitter(img,nrow(img)/10,ncol(img)/10) #Split the img into 100 sub-images
str(split_img)
```
Now that we've split our image into sub-images, can we get the mean of each sub-image and plot it?
```{r}
mean(split_img[,,1]) # get mean of first sub-image
subimagemean <- apply(split_img,3,mean) # get mean of all sub-images, averaging over 1st and 2nd dimensions
hist(subimagemean) #put it in a histogram
ggplot() + aes(subimagemean)+ geom_histogram(binwidth=200, colour="black", fill="white") #use ggplot
```
How well does our sampling compare to the true mean? And how many times do we have to subsample to get an accurate reading of the true mean?
```{r}
ggplot() + 
  aes(subimagemean)+ 
  geom_histogram(binwidth=200, colour="black", fill="white")+
  geom_vline(aes(xintercept=mean(img)),
            color="blue", linetype="dashed", size=1)
# Note that mean(img)=mean(subimagemean), following the central limit theorem.
# What is the variance of our sampling distribution?
var(subimagemean) # variance of sampling distribution
sd(img)^2 # variance of original img
sd(img)^2/var(subimagemean)   # show central limit theorem
```


